Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics

### 2021-04-14 various_stuff/dev_vignettes/inhibition strength study.Rmd
reorganisation of vignettes
now need to get them working!

---
title: "Diversity and the nature of responses to environmental change."
author: "Owen Petchey & Jennifer Probst"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 4
    number_sections: true
    code_folding: hide
    toc_float:
      collapsed: true
      smooth_scroll: true
---

# Setup R

```{r setup}
rm(list=ls())
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

library(tidyverse)
library(deSolve)
library(rootSolve)

set.seed(13)

run_sims <- FALSE
```

@@

* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).
* Add an investigation of how changes in tolerances alter the response to environmental change.

We also add:

* Option to include random noise in substrate concentrations.
* Option to temporally vary oxygen diffusivity.
* Option to include immigration.
* Option to include a minimum abundances.

Here we use only the published (Bush et al 2017) dynamical model, in which there is not biodiversity within each of the three functional groups. That is, there is only one parameter set per functional group.

Here is the original publication:
https://www.nature.com/articles/s41467-017-00912-x

And here is the supplementary information, including the table of parameter values:
https://static-content.springer.com/esm/art%3A10.1038%2Fs41467-017-00912-x/MediaObjects/41467_2017_912_MOESM1_ESM.pdf

Note that we here use the notation from the table S1 of the supplementary information, and not the notation in the main text. (There is only one small difference between these: In the main text the notation is simplified to y_S_PB (which is y_SR_PB ($y_{SR,PB}$) in Table S1) and y_S_SB (which is y_SO_SB ($y_{SO,SB}$) in Table S1).

@@

## Model dimensions

* Time: hours
* Volume: litres
* Substrate quantity: micromoles
* Organism quantity: cells

## Dynamic model

```{r}
source("functions/bushplus_dynamic_model.r")
```

## Defaults

The framework used for running a simulation, and for varying aspects of a simulation run is:

* Define the default simulation conditions.
* Use these defaults unless other conditions are specified.
* The function call `run_simulation()` would therefore run a simulation with default conditions.
* Whereas the function call `run_simulation(initial_state = init_state_favouring_anoxic)` would run the simulation with specified initial state values.

@@

# Results: Reproducing previous results

## Initial conditions favouring anoxic, no noise, no forcing

Figure 2a, b in Bush et al.

```{r}

init_state_favouring_anoxic <- c(N_CB = 5e1,
           N_PB = 1e7,
           N_SB = 1e7,
           SO = 300,
           SR = 300,
           O = 1e1,
           P = 1e1)

sim_res <- run_simulation(initial_state = init_state_favouring_anoxic)
plot_dynamics(simulation_result = sim_res)

#ggsave('bacteria_no_noise.png')
```

Population dynamics seem as in the paper.
Resource dynamics fluctuate more than in the paper.

## Initial conditions favouring anoxic, with noise, no forcing

Figure 2a, b in Bush et al, but with noise

```{r}
## add some noise
noise_sigma_0.01 <- 0.01
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          noise_sigma = noise_sigma_0.01)
plot_dynamics(sim_res)
#ggsave("high_SB_PB.png")
```

## Initial conditions favouring oxic, no noise, no forcing

Figure 2c, d in Bush et al.

```{r} 
init_state_favouring_oxic <- c(N_CB = 1e7,
           N_PB = 1e5,
           N_SB = 1e5,
           SO = 500,
           SR = 50,
           O = 30,
           P = 10)

sim_res <- run_simulation(initial_state = init_state_favouring_oxic)

plot_dynamics(sim_res)
##ggsave("high_CB.png")
```
Population and resource dynamics seen as those in the paper.

## Initial conditions favouring oxic, with noise, no forcing

Figure 2c, d in Bush et al. but with noise.

```{r}

sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          noise_sigma = noise_sigma_0.01)

plot_dynamics(sim_res)
##ggsave("high_CB.png")
```

# Results: Composition/diversity and nature of response

```{r}
ssfind_minimum_abundances <- rep(0, 3)
names(ssfind_minimum_abundances) <- c("CB", "PB", "SB") 
ssfind_simulation_duration <- 10000
ssfind_simulation_sampling_interval <- ssfind_simulation_duration
ssfind_event_interval <- ssfind_simulation_duration
ssfind_parameters <- default_parameter_values
# ssfind_init_state <- c(N_CB = NA, 
#                        N_PB = NA,
#                        N_SB = NA,
#                        SO = 500,
#                        SR = 50,
#                        O = 30,
#                        P = 10)
##ssfind_init_state <- default_initial_state
ssfind_init_state <- init_state_favouring_anoxic
ssfind_init_state <- c(N_CB = NA, #this one is varied
           N_PB = 1e8,
           N_SB = 1e8,
           SO = 200,
           SR = 200,
           O = 10,
           P = 9.5)
grid_num_a <- 100
a_Os <- 10^seq(-4.5, -1.5, length=grid_num_a)
grid_num_N <- 2
```

### Each composition

#### No organisms

```{r eval = run_sims}

initial_CBs <- 0
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_no_biology <- ss_by_a_N(expt)
saveRDS(ss_no_biology, "data/ss_no_biology.RDS")
```

```{r}
ss_no_biology <- readRDS("data/ss_no_biology.RDS")
```

```{r}
ss_no_biology %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_no_biology %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_CB <- ss_by_a_N(expt)
saveRDS(ss_only_CB, "data/ss_only_CB.RDS")
```

```{r}
ss_only_CB <- readRDS("data/ss_only_CB.RDS")
```

```{r}
tempx1 <- ss_only_CB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  geom_path(data = filter(tempx1, Species == "PB"),
            aes(y=0.2), size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only CB")

ggsave("figures/CB.png", width = 3, height = 2)
```

```{r}
ss_only_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 0
initial_SBs <- 10^seq(0, 8, length=grid_num_N)
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_SB <- ss_by_a_N(expt)

saveRDS(ss_only_SB, "data/ss_only_SB.RDS")
```

```{r}
ss_only_SB <- readRDS("data/ss_only_SB.RDS")
```

```{r}
ss_only_SB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_only_PB <- ss_by_a_N(expt)

saveRDS(ss_only_PB, "data/ss_only_PB.RDS")
```

```{r}
ss_only_PB <- readRDS("data/ss_only_PB.RDS")
```

```{r}
tempx1 <- ss_only_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.))
tempx1 %>%
  ggplot(aes(x = a, y = log10(Quantity), col = Species)) +
  geom_path(size = 0.8) +
  theme_light() +
  ylim(0,10) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  geom_path(data = filter(tempx1, Species == "CB"),
            aes(y=0.2), size = 0.8)+
  geom_path(data = filter(tempx1, Species == "PB"),
            size = 0.8) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="Only PB")

ggsave("figures/PB.png", width = 3, height = 2)

```{r}
ss_only_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_PB

ss_only_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_PB, "data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB <- readRDS("data/ss_only_SB_PB.RDS")
```

```{r}
ss_only_SB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### SB and CB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 0
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_SB <- expt$N_CB

ss_only_SB_CB <- ss_by_a_N(expt)
saveRDS(ss_only_SB_CB, "data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB <- readRDS("data/ss_only_SB_CB.RDS")
```

```{r}
ss_only_SB_CB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_CB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:5) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB and PB

```{r eval = run_sims}

initial_CBs <- 0  # this one is varied
initial_PBs <- 10^seq(0, 8, length=grid_num_N)
initial_SBs <- 0
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)
expt$N_CB <- expt$N_PB

ss_only_CB_PB <- ss_by_a_N(expt)
saveRDS(ss_only_CB_PB, "data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB <- readRDS("data/ss_only_CB_PB.RDS")
```

```{r}
ss_only_CB_PB %>%
  select(a, N_CB, N_SB, N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

```{r}
ss_only_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

#### CB, SB, and PB

```{r eval = run_sims}

initial_CBs <- 10^seq(0, 8, length=grid_num_N)
initial_PBs <- 1e5
initial_SBs <- 1e5
expt <- expand.grid(N_CB = initial_CBs,
                    N_PB = initial_PBs,
                    N_SB = initial_SBs,
                    a_O = a_Os)

ss_CB_SB_PB <- ss_by_a_N(expt)
saveRDS(ss_CB_SB_PB, "data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB <- readRDS("data/ss_CB_SB_PB.RDS")
```

```{r}
ss_CB_SB_PB %>%
  select(a, CB=N_CB, SB=N_SB, PB=N_PB) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species),
            size=0.8) +
  ylab(expression(Log[10](Abundance))) +
  xlab(expression(Log[10](Oxygen~diffusivity))) +
  theme_light() +
  ylim(0,10) +
  theme(plot.title=element_text(size=10, hjust=0.5)) +
  labs(title="CB, SB, & PB") 
ggsave("figures/CB_SB_PB.png", width = 3, height = 2)

```{r}
ss_CB_SB_PB %>%
  select(a, SO, SR, O, P) %>%
  gather(key = Species, value=Quantity, 2:ncol(.)) %>%
  ggplot() +
  geom_path(aes(x = a, y = log10(Quantity), col = Species))
```

### All response measures

```{r eval = run_sims}
all_ss <- bind_cols(readRDS("data/ss_no_biology.RDS"),
                    composition = rep("No organisms", grid_num_a*grid_num_N),
                    direction = c(rep("up", grid_num_a),
                                  rep("down", grid_num_a)))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB.RDS"),
                              composition = rep("CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB.RDS"),
                              composition = rep("SB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_PB.RDS"),
                              composition = rep("PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_PB.RDS"),
                              composition = rep("SB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_SB_CB.RDS"),
                              composition = rep("SB-CB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_only_CB_PB.RDS"),
                              composition = rep("CB-PB", grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))
all_ss <- bind_rows(all_ss,
                    bind_cols(readRDS("data/ss_CB_SB_PB.RDS"),
                              composition = rep("CB-SB-PB",
                                                grid_num_a*grid_num_N),
                              direction = c(rep("up", grid_num_a),
                                            rep("down", grid_num_a))))

x_long <- all_ss %>%
  gather(key = Species, value = Quantity, 2:8)
x_wide <- x_long %>%
  select(-initial_N_CB) %>%
  spread(key = direction, value=Quantity, drop=T)

resp_summary <- x_wide %>%
  group_by(composition, Species) %>%
  summarise(hyst_1 = mean(abs(log10(up) - log10(down))),
            hyst_2 = get_hysteresis_range(up, down, a),
            nl_up = get_nonlinearity(a, log10(up)),
            nl_down = get_nonlinearity(a, log10(down))) %>%
  ungroup() %>%
  gather(key = resp_measure, value = value, 3:6) %>%
  mutate(composition = fct_relevel(composition,
                                   "No organisms",
                                   "CB",
                                   "SB",
                                   "PB",
                                   "CB-PB",
                                   "SB-CB",
                                   "SB-PB",
                                   "CB-SB-PB"),
         num_func_groups = case_when(composition =="No organisms" ~ 0,
                                     composition == "CB" ~ 1,
                                     composition == "SB" ~ 1,
                                     composition == "PB" ~ 1,
                                     composition == "CB-PB" ~ 2,
                                     composition == "SB-CB" ~ 2,
                                     composition == "SB-PB" ~ 2,
                                     composition == "CB-SB-PB" ~ 3
                                     ),
         var_type = ifelse(str_sub(Species, 1, 1)=="N", "Organism", "Substrate")
         )
saveRDS(resp_summary, "data/resp_summary.RDS")
```

```{r}
resp_summary <- readRDS("data/resp_summary.RDS")
resp_summary %>%
  ggplot() +
  geom_point(aes(x = composition, y = value)) +
  facet_grid(Species ~ resp_measure, scales = "free") +
  coord_flip()

```{r}
resp_summ1 <- resp_summary %>%
  filter(resp_measure %in% c("nl_down", "nl_up")) %>%
  group_by(composition, Species, num_func_groups, var_type) %>%
  summarise(ave_nl = mean(value))

resp_summ1 %>%
  ggplot() +
  geom_point(aes(composition, ave_nl)) +
  facet_wrap( ~ Species)

```{r}
focus_on <- "Substrate"
ave_nl_substrate <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = mean(ave_nl))
ave_nl_substrate %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

```{r}
focus_on <- "Organism"
ave_nl_organism <- resp_summ1 %>%
  filter(var_type == focus_on) %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl)/unique(num_func_groups))
ave_nl_organism %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
```

```{r}
focus_on <- "System"
ave_nl_system <- resp_summ1 %>%
  group_by(composition) %>%
  summarise(ave_ave_nl = sum(ave_nl) / (unique(num_func_groups)+4))
ave_nl_system %>%
  ggplot() +
  geom_point(aes(composition, ave_ave_nl)) +
  ggtitle(paste("Nonlinearity in", focus_on, "response."))
ggsave("figures/system_nonlinearity.png", width = 3, height = 2)

```{r}
focus_on <- "Substrate"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = mean(value))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```{r}
focus_on <- "Organism"
resp_summ3 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2"),
         var_type == focus_on) %>%
  group_by(composition, num_func_groups, resp_measure, var_type) %>%
  summarise(ave_hyst = sum(value) / unique(num_func_groups))

resp_summ3 %>%
  ggplot() +
  geom_point(aes(x = composition, y = ave_hyst)) +
  facet_wrap( ~ resp_measure, scales="free") +
  ggtitle(paste("Hysteresis in", focus_on, "response."))

```{r}
resp_summ4 <- resp_summary %>%
  filter(resp_measure %in% c("hyst_1", "hyst_2")) %>%
  group_by(resp_measure, composition) %>%
  summarise(mean_hyst = sum(value) / (unique(num_func_groups)+4))

resp_summ4 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_hyst)) +
  facet_wrap( ~ resp_measure) +
  ggtitle("Hysteresis in system response.")

```{r}
resp_summ5 <- resp_summ4 %>%
  group_by(composition) %>%
  summarise(mean_mean_hyst = mean(mean_hyst))

resp_summ5 %>%
  ggplot() +
  geom_point(aes(x = composition, y = mean_mean_hyst)) +
  xlab("Functional groups present") +
  ylab("Amount of hysteresis") +
  ylim(0,2) +
  theme_light() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
ggsave("figures/composition_system_hysteresis.png", width = 2.5, height = 2.5)

@@

# Results: Temporal state switching

## Switch anoxic-oxic-anoxic

```{r eval = run_sims}
## anoxic initial state
init_state <- init_state_favouring_anoxic

## simulation timing
simulation_duration <- 5000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
lhl_log10a_series <- c(-8, -8, -2, -2, -2, -2, -2, -8, -8)

## run simulation
sim_res <- run_simulation(initial_state = init_state_favouring_anoxic,
                          log10a_series = lhl_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)
saveRDS(sim_res, "data/lhl_with_defaults.RDS")
```

```{r}
lhl_with_defaults <- readRDS("data/lhl_with_defaults.RDS")
plot_dynamics(lhl_with_defaults)
```

```{r eval = run_sims}
lhl_with_defaults$result %>%
  filter(between(time, 1000, 73000)) %>%
  ggplot() +
  geom_path(aes(x=a, y=log10(O)),
            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
```

## Switch oxic-anoxic-oxic

```{r eval = run_sims}
## anoxic initial state
init_state <- init_state_favouring_oxic

## simulation timing
simulation_duration <- 50000
simulation_sampling_interval <- 10
event_interval_t <- 100

## low-high-low oxgygen diffusivity 
hlh_log10a_series <- c(-2, -2, -10, -14, -18, -22, -2, -2, -2)

## run simulation
sim_res <- run_simulation(initial_state = init_state_favouring_oxic,
                          log10a_series = hlh_log10a_series,
                          sim_duration = simulation_duration,
                          sim_sample_interval = simulation_sampling_interval,
                          event_interval = event_interval_t)
saveRDS(sim_res, "data/hlh_with_defaults.RDS")
```

```{r}
hlh_with_defaults <- readRDS("data/hlh_with_defaults.RDS")
plot_dynamics(hlh_with_defaults)
```

```{r eval = run_sims}
hlh_with_defaults$result %>%
  filter(between(time, 1000, 100000)) %>%
  ggplot() +
  geom_path(aes(x=a, y=log10(O)),
            arrow = arrow(type = "open", angle = 30, length = unit(0.1, "inches")))
```

@@

## EWS analyses

Evaluation of the switch: 
Get the functions: 
```{r eval = F}
source('C:/Users/probs/Desktop/Sem 6/Internship/functions_analysis_EWS_2.R')
```

Get state shifts and calculate parameters: 
```{r eval = F}
x3 <- 30 #size of wished rolling window (percentage btw 0 and 100)
x5 <- 50 #size of wished rolling window (percentage btw 0 and 100)

#first switch
name_1='anoxic_oxic 1'
trans1 <- data.frame(out_t %>% filter(time>87500 & time<92500)) #for analysis
trans1_f <- data.frame(out_f %>% filter(time>87500 & time<92500)) #for plotting
n_13 <- round(length(trans1$time)*x3/100) #size of rolling window
t_13 <- length(trans1$time) - n_13 + 1 #number of rolling windows
n_15 <- round(length(trans1$time)*x5/100) #size of rolling window
t_15 <- length(trans1$time) - n_15 + 1 #number of rolling windows
out_df_1_N_CB <- data.frame(trans1$N_CB)
colnames(out_df_1_N_CB) <- c("N_CB")
out_df_1_N_PB <- data.frame(trans1$N_PB)
colnames(out_df_1_N_PB) <- c("N_PB")
out_df_1_N_SB <- data.frame(trans1$N_SB)
colnames(out_df_1_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans1_f) +
  ggtitle("Anoxic_oxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 1.png")

#second switch
name_2='oxic_anoxic 1'
trans2 <- data.frame(out_t %>% filter(time>335000 & time<360000))
trans2_f <- data.frame(out_f %>% filter(time>335000 & time<360000))
n_23 <- round(length(trans2$time)*x3/100) 
t_23 <- length(trans2$time) - n_23 + 1
n_25 <- round(length(trans2$time)*x5/100) 
t_25 <- length(trans2$time) - n_25 + 1
out_df_2_N_CB <- data.frame(trans2$N_CB)
colnames(out_df_2_N_CB) <- c("N_CB")
out_df_2_N_PB <- data.frame(trans2$N_PB)
colnames(out_df_2_N_PB) <- c("N_PB")
out_df_2_N_SB <- data.frame(trans2$N_SB)
colnames(out_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans2_f) +
  ggtitle("Oxic_anoxic 1")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 1.png")

#third switch
name_3='oxic_anoxic 2'
trans3 <- data.frame(out2_t %>% filter(time>165000 & time<185000))
trans3_f <- data.frame(out2_f %>% filter(time>165000 & time<185000))
n_33 <- round(length(trans3$time)*x3/100) 
t_33 <- length(trans3$time) - n_33 + 1
n_35 <- round(length(trans3$time)*x5/100) 
t_35 <- length(trans3$time) - n_35 + 1
out2_df_1_N_CB <- data.frame(trans3$N_CB)
colnames(out2_df_1_N_CB) <- c("N_CB")
out2_df_1_N_SB <- data.frame(trans3$N_SB)
colnames(out2_df_1_N_SB) <- c("N_SB")
out2_df_1_N_PB <- data.frame(trans3$N_PB)
colnames(out2_df_1_N_PB) <- c("N_PB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans3_f) +
  ggtitle("Oxic_anoxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("oxic_anoxic 2.png")

#fourth switch
name_4='anoxic_oxic 2'
trans4 <- data.frame(out2_t %>% filter(time>340000 & time<348500))
trans4_f <- data.frame(out2_f %>% filter(time>340000 & time<348500))
n_43 <- round(length(trans4$time)*x3/100) 
t_43 <- length(trans4$time) - n_43 + 1
n_45 <- round(length(trans4$time)*x5/100) 
t_45 <- length(trans4$time) - n_45 + 1
out2_df_2_N_CB <- data.frame(trans4$N_CB)
colnames(out2_df_2_N_CB) <- c("N_CB")
out2_df_2_N_PB <- data.frame(trans4$N_PB)
colnames(out2_df_2_N_PB) <- c("N_PB")
out2_df_2_N_SB <- data.frame(trans4$N_SB)
colnames(out2_df_2_N_SB) <- c("N_SB")

ggplot(aes(x=time, y=trans_quantity, col=species), data=trans4_f) +
  ggtitle("Anoxic_oxic 2")+
  xlab('Time') +
  ylab('Trans_quantity') +
  facet_wrap(~var_type, scales = "free", ncol = 1) + #create one plot for organisms and one for substrate
  geom_line()+
  labs(color = "Organism/Substrate")
ggsave("anoxic_oxic 2.png")
```

Rolling window analysis
```{r eval = F}
#analysis for 1st
rollingwindow_analysis(trans1, 2, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 3, n_15, t_15, name_1)
rollingwindow_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
rollingwindow_analysis(trans2, 2, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 3, n_25, t_25, name_2)
rollingwindow_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
rollingwindow_analysis(trans3, 2, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 3, n_35, t_35, name_3)
rollingwindow_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
rollingwindow_analysis(trans4, 2, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 3, n_45, t_45, name_4)
rollingwindow_analysis(trans4, 4, n_45, t_45, name_4)

Metric based indicators from the earlywarnings package
```{r eval = F}
#analysis for 1st
earlywarnings_plot(out_df_1_N_CB, name_1, x3)
earlywarnings_plot(out_df_1_N_PB, name_1, x3)
earlywarnings_plot(out_df_1_N_SB, name_1, x3)

#analysis for 2nd
earlywarnings_plot(out_df_2_N_CB, name_2, x3)
earlywarnings_plot(out_df_2_N_PB, name_2, x3)
earlywarnings_plot(out_df_2_N_SB, name_2, x3)

#analysis for 3rd
earlywarnings_plot(out2_df_1_N_CB, name_3, x3)
earlywarnings_plot(out2_df_1_N_PB, name_3, x3)
earlywarnings_plot(out2_df_1_N_SB, name_3, x3)

#analysis for 4th
earlywarnings_plot(out2_df_2_N_CB, name_4, x3)
earlywarnings_plot(out2_df_2_N_PB, name_4, x3)
earlywarnings_plot(out2_df_2_N_SB, name_4, x3)
```

Conditional Heteroskedasticity 
```{r eval = F}
#analysis for 1st
ch <- condhet_analysis(out_df_1_N_CB, name_1)
ch <- condhet_analysis(out_df_1_N_PB, name_1)
ch <- condhet_analysis(out_df_1_N_SB, name_1)

#analysis for 2nd
ch2 <- condhet_analysis(out_df_2_N_CB, name_2)
ch2 <- condhet_analysis(out_df_2_N_PB, name_2)
ch2 <- condhet_analysis(out_df_2_N_SB, name_2)

#analysis for 3rd
ch3 <- condhet_analysis(out2_df_1_N_CB, name_3)
ch3 <- condhet_analysis(out2_df_1_N_PB, name_3)
ch3 <- condhet_analysis(out2_df_1_N_SB, name_3)

#analysis for 4th
ch4 <- condhet_analysis(out2_df_2_N_CB,name_4)
ch4 <- condhet_analysis(out2_df_2_N_PB, name_4)
ch4 <- condhet_analysis(out2_df_2_N_SB, name_4)
```

Detrending fluctuation analysis 
```{r eval = F}
#analysis for 1st
dfa_analysis(trans1, 2, n_15, t_15, name_1)
dfa_analysis(trans1, 3, n_15, t_15, name_1)
dfa_analysis(trans1, 4, n_15, t_15, name_1)

#analysis for 2nd
dfa_analysis(trans2, 2, n_25, t_25, name_2)
dfa_analysis(trans2, 3, n_25, t_25, name_2)
dfa_analysis(trans2, 4, n_25, t_25, name_2)

#analysis for 3rd
dfa_analysis(trans3, 2, n_35, t_35, name_3)
dfa_analysis(trans3, 3, n_35, t_35, name_3)
dfa_analysis(trans3, 4, n_35, t_35, name_3)

#analysis for 4th
dfa_analysis(trans4, 2, n_45, t_45, name_4)
dfa_analysis(trans4, 3, n_45, t_45, name_4)
dfa_analysis(trans4, 4, n_45, t_45, name_4)
```

Drift Diffusion Jump Nonparametrics Early Warning Signals:
```{r eval = F}
#shift 1
ddj11<- ddj_model(out_df_1_N_CB)
plot_ddj_1(out_df_1_N_CB, ddj11, name_1)
plot_ddj_2(out_df_1_N_CB, ddj11, name_1)

>>>>

---
title: "Partial reproduction of Bush et al (2017) Oxic-anoxic regime shifts mediated by feedbacks between biogeochemical processes and microbial community dynamics"
author: "Owen Petchey"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Reproducing some results of Bush et al 2017}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r setup}
## Clear R
rm(list=ls())

## some chunk options
knitr::opts_chunk$set(echo = TRUE, cache = TRUE)

## load required libraries
library(tidyverse)
library(deSolve)
library(rootSolve)
library(microxanox)

## Set random number generator seed
## Is redundant if there are no stochastic components in the model
set.seed(13)

## Should the rather lengthy simulations required to 
## construct the state-environment relationship
## be run?
run_steadystate_sims <- FALSE
```

@@

* Reproduce the results of [Bush et al 2017](https://www.nature.com/articles/s41467-017-00912-x).

Please see the User manual vignette for further introduction and general methods.

@@

@@

@@

@@
